Parallel Hierarchical Clustering using \\ Rank-Two Nonnegative Matrix Factorization \thanks{Identify applicable funding agency here. If none, delete this.}

### 2020-07-16 latex/paper.tex
replaced IMAGE with macro

\GB{need shortcut names for datasets: synthetic, DC, and something better than IMAGE}

\begin{itemize}
	\item \textit{Hyperspectral Imaging}:
	\LM{DC: https://engineering.purdue.edu/~biehl/MultiSpec/hyperspectral.html}
	For Hyperspectral imaging, we chose the Hyperspectral digital imagery collection experiment (HYDICE) image of the Washington DC Mall. We will refer
	to this dataset as \hyper{}.
	\hyper{} is formatted into a 3-way tensor representing two spatial dimensions of pixels and one dimension of spectral bands. So, a slice along
	the spectral band dimnesion would be the full \hyper{} image in that spectral band. For H2NMF, these tensors are flattened so that the rows represent the
	$191$ spectral bands and the columns represent the $392960$ pixels. 
	Larger hyperspectral datasets tend to not grow larger than around 200 spectral bands and have millions or billions of pixels. So, scaling is 
	limited for this dataset given our data distribution scheme.
	\item \textit{Image Classification}:
	\LM{https://www.kaggle.com/c/siim-isic-melanoma-classification/data}
	Image Classification datasets contain tens of thousands of images of equal size. The dataset we chose is the SIIM-ISIC Melanoma classification
	dataset, which we will refer to as \image{}. \image{} consists of $33126$ RGB training images equally sized at $1024 \times 1024$. Unlike with hyperspectral imaging, the resulting
	matrix used in H2NMF consists of image pixels along the rows and individual images along the columns. So, the resulting sized matrix is
	$3145728 \times 33126$, which is approximately 800GB in size. 
	Given its size, \image{} can only scale from $10$ Summit nodes.
	\item \textit{Synthetic Dataset}:
	Our synthetic dataset is the same shape of \image{} but consists of fewer rows and columns by a factor of $3$. The resulting matrix is $1048576 \times 11042$. 
	We chose a smaller size dataset to better see the limits of scaling and to scale from fewer Summit nodes.
\end{itemize}

@@

We perform strong scaling experiments for a single Rank-2 NMF (\Cref{alg:parrank2nmf}) on the synthetic and IMAGE datasets.
The theory (\Cref{eq:r2nmfcost}) suggests that perfect strong scaling is possible as long as the execution time is dominated by local computation.
Both the matrix multiplications and NNLS solves scale linearly with $1/p$ (we expect MatMul to dominate), but the bandwidth cost is independent of $p$ and the latency cost increases slightly with $p$.

\Cref{fig:synrank2speedup,fig:rwrank2speedup} show performance relative to the smallest number of nodes required to store data and factor matrices.
The synthetic data is chosen to fit on a single node, while the IMAGE data requires 10 nodes to store the input matrix as well as temporary and output matrices.
For these data sets, we observe nearly perfect strong scaling, with 47$\times$ speedup on 50 nodes (over 1 node) for synthetic data and 7.1$\times$ speedup on 80 nodes (over 10 nodes) for IMAGE data.

@@

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_rank2_speedup.pdf}
\caption{Strong Scaling Speedup for Rank-2 NMF on IMAGE data}
\label{fig:rwrank2speedup}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_rank2_strongscaling.pdf}
\caption{Relative Time Breakdown for Rank-2 NMF on IMAGE data}
\label{fig:rwrank2strongscaling}
\end{center}
\end{figure}

>>>>

\GB{need shortcut names for datasets: synthetic, DC, and something better than \image{}}

\begin{itemize}
	\item \textit{Hyperspectral Imaging}:
	\LM{DC: https://engineering.purdue.edu/~biehl/MultiSpec/hyperspectral.html}
	For Hyperspectral imaging, we chose the Hyperspectral digital imagery collection experiment (HYDICE) image of the Washington DC Mall. We will refer
	to this dataset as \hyper{}.

	\hyper{} is formatted into a 3-way tensor representing two spatial dimensions of pixels and one dimension of spectral bands. So, a slice along
	the spectral band dimnesion would be the full \hyper{} image in that spectral band. For H2NMF, these tensors are flattened so that the rows represent the
	$191$ spectral bands and the columns represent the $392960$ pixels.

	Larger hyperspectral datasets tend to not grow larger than around 200 spectral bands and have millions or billions of pixels. So, scaling is 
	limited for this dataset given our data distribution scheme.
	\item \textit{Image Classification}:
	\LM{https://www.kaggle.com/c/siim-isic-melanoma-classification/data}
	Image Classification datasets contain tens of thousands of images of equal size. The dataset we chose is the SIIM-ISIC Melanoma classification
	dataset, which we will refer to as \image{}. \image{} consists of $33126$ RGB training images equally sized at $1024 \times 1024$. Unlike with hyperspectral imaging, the resulting
	matrix used in H2NMF consists of image pixels along the rows and individual images along the columns. So, the resulting sized matrix is
	$3145728 \times 33126$, which is approximately 800GB in size.

	Given its size, \image{} can only scale from $10$ Summit nodes.
	\item \textit{Synthetic Dataset}:
	Our synthetic dataset is the same shape of \image{} but consists of fewer rows and columns by a factor of $3$. The resulting matrix is $1048576 \times 11042$. 
	We chose a smaller size dataset to better see the limits of scaling and to scale from fewer Summit nodes.
\end{itemize}

@@

We perform strong scaling experiments for a single Rank-2 NMF (\Cref{alg:parrank2nmf}) on the synthetic and \image{} datasets.
The theory (\Cref{eq:r2nmfcost}) suggests that perfect strong scaling is possible as long as the execution time is dominated by local computation.
Both the matrix multiplications and NNLS solves scale linearly with $1/p$ (we expect MatMul to dominate), but the bandwidth cost is independent of $p$ and the latency cost increases slightly with $p$.

\Cref{fig:synrank2speedup,fig:rwrank2speedup} show performance relative to the smallest number of nodes required to store data and factor matrices.
The synthetic data is chosen to fit on a single node, while the \image{} data requires 10 nodes to store the input matrix as well as temporary and output matrices.
For these data sets, we observe nearly perfect strong scaling, with 47$\times$ speedup on 50 nodes (over 1 node) for synthetic data and 7.1$\times$ speedup on 80 nodes (over 10 nodes) for \image{} data.

@@

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_rank2_speedup.pdf}
\caption{Strong Scaling Speedup for Rank-2 NMF on \image{} data}
\label{fig:rwrank2speedup}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_rank2_strongscaling.pdf}
\caption{Relative Time Breakdown for Rank-2 NMF on \image{} data}
\label{fig:rwrank2strongscaling}
\end{center}
\end{figure}
